Numerical simulation and entropy dissipative cure of the carbuncle instability for the shallow water circular hydraulic jump

### 2021-02-21 numerics_paper/version2/numerical_chj.tex

\begin{figure}[!h]
  \centering
  \subfloat[Solution at (left) $t=1$, (middle) $t=1.5$ and (right) $t=3$ using Roe's Riemann solver.
    The main plots are Schlieren plots of the depth $h$, while the inset in the right
    figure shows the magnitude of the momentum.\label{fig:chj_r1_roe}]{
    \begin{tabular}{ccc} 
      \includegraphics[scale=0.3]{figures/chj_r1_roe_p1.png} &
      \includegraphics[scale=0.3]{figures/chj_r1_roe_p2.png} &
      \includegraphics[scale=0.3]{figures/chj_r1_roe_p3.png}
    \end{tabular}
  }

@@

    \caption{Simulation of a CHJ with boundary conditions given by \eqref{bcs_chj} and \eqref{bcs_chj_r1}.    
      We consider method \eqref{second-order_via_fluct} with different Riemann solvers.
      In all cases we show the Schlieren plot for the water height $h$.
      In (a), for the plot that corresponds to $t=3$, we zoom close to the jump
      and plot the y-momentum. 
    \label{fig:chj_r1}}
\end{figure}

@@

  \vspace{10pt}
  \subfloat[Solution at $t=0.1$ using (left) Rusanov's and (right) the blended solvers.\label{fig:chj_r2_diff}]{
    \begin{tabular}{ccc}
      \includegraphics[scale=0.3]{figures/chj_r2_llf_t0p1.png} \qquad &
      %\quad \includegraphics[scale=0.5]{figures/cross.png} \quad &
      \qquad \includegraphics[scale=0.3]{figures/chj_r2_blended_t0p1.png}
    \end{tabular}
    }
  \caption{CHJ with boundary conditions given by \eqref{bcs_chj} and \eqref{bcs_chj_r2}. 
    We consider method \eqref{second-order_via_fluct} with different Riemann solvers.
    In all cases we show the Schlieren plot for the water height $h$.
    In (a), for the plot that corresponds to $t=0.11$, we zoom close to the jump and plot the y-momentum.}
\end{figure}

\begin{figure}[!h]
  \centering
    \includegraphics[scale=0.3]{figures/chj_r2_rusanov_t5p0.png} \quad
    %\includegraphics[scale=0.5]{figures/cross.png} \quad
    \quad \includegraphics[scale=0.3]{figures/chj_r2_blended_t5p0.png}
  \caption{CHJ at $t=5$ with boundary conditions given by \eqref{bcs_chj} and \eqref{bcs_chj_r2}. 
    We consider method \eqref{second-order_via_fluct} with (left) Rusanov's
    and (right) the blended solvers. 
    In all cases we superimpose the Schlieren plot for the water height $h$
    an the magnitude of the momentum with a uniform (white-to-blue) colormap. 
    The scale for colormap is (white) $0.45$ to (blue) $4.5$, 
    which corresponds to the minimum and maximum values at $t=0$.
    \label{fig:chj_r2_later}}
\end{figure}

\subsubsection{Regime II with random perturbations}\label{sec:regime_iiwp}
Given the symmetry of the grid and initial data in the section above,
the non-rotationally-symmetric perturbations seen when using the blended
solver above must arise due to the influence of numerical errors.
In order to understand whether these are evidence of a true instability,
we now introduce a non-symmetric perturbation at the inflow boundary.
We use the same mean values \eqref{bcs_chj_r2}, but now we set
\begin{subequations}
  \begin{align}
    h(x,y,t)=\tilde h_\jet, \quad
    u(x,y,t)=|\tilde \bfu_\jet| \left(\frac{x}{r_\jet}\right), \quad
    v(x,y,t)=|\tilde \bfu_\jet| \left(\frac{y}{r_\jet}\right), \qquad r=r_\jet,
  \end{align}
\end{subequations}
with $\tilde h_\jet = \frac{h_\jet}{1+\epsilon(x,y,t)}$ and $|\tilde\bfu_\jet|=|\bfu_\jet|(1+\epsilon(x,y,t))$.
Here $\epsilon(x,y,t)$ is chosen at each inflow boundary ghost cell and at each
time step as an i.i.d. random variable from the uniform distribution $[-0.01,0.01]$.
In this case, due to the large Froude number, the random perturbations in the inflow
may trigger physical instabilities, which we want to preserve. At the same time, we
do not want the solution to develop carbuncle instabilities.
In figures \ref{fig:chj_r2wp_rusanov} and \ref{fig:chj_r2wp_blended}, we show the solution
at different times considering method \eqref{second-order_via_fluct} with Rusanov's and
the blended Riemann solvers, respectively.
In the right panel of Figure \ref{fig:chj_r2wp_blended}, we show an inset figure 
with the y-momentum. Note that even though the solution is highly unstable 
(with the blended Riemann solver), no visible carbuncle instabilities are developed.

@@

  \subfloat[Blended Riemann solver\label{fig:chj_r2wp_blended}]{
    \begin{tabular}{cccc}
      \includegraphics[scale=0.24]{figures/chj_r2wp_blended_t2p0.png} &
      \includegraphics[scale=0.24]{figures/chj_r2wp_blended_t3p0.png} &
      \includegraphics[scale=0.24]{figures/chj_r2wp_blended_t4p0.png} &
      \includegraphics[scale=0.24]{figures/chj_r2wp_blended_t5p0.png}
    \end{tabular}
  }
  \caption{CHJ at different times with a random perturbation at the inflow boundary.
    We consider method \eqref{second-order_via_fluct} with different Riemann solvers. 
    In all cases we show the Schlieren plot for the water height $h$.
    In (b), for the plot that corresponds to $t=5$, we zoom close to the jump and plot the y-momentum.}
\end{figure}

>>>>

\begin{figure}[!h]
  \centering
  \subfloat[Solution at (left) $t=1$, (middle) $t=1.5$ and (right) $t=3$ using Roe's Riemann solver.
    The main plots are Schlieren plots of the depth $h$, while the inset in the right
    figure shows the momentum in the radial direction.\label{fig:chj_r1_roe}]{
    \begin{tabular}{ccc} 
      \includegraphics[scale=0.3]{figures/chj_r1_roe_p1.png} &
      \includegraphics[scale=0.3]{figures/chj_r1_roe_p2.png} &
      \includegraphics[scale=0.3]{figures/chj_r1_roe_p3.png}
    \end{tabular}
  }

@@

    \caption{Simulation of a CHJ with boundary conditions given by \eqref{bcs_chj} and \eqref{bcs_chj_r1}.    
      We consider method \eqref{second-order_via_fluct} with different Riemann solvers.
      In (a) and (b) we show the Schlieren plot for the water height $h$.
    \label{fig:chj_r1}}
\end{figure}

@@

  \vspace{10pt}
  \subfloat[Solution at $t=0.11$ using (left) Rusanov's and (right) the blended solvers.\label{fig:chj_r2_diff}]{
    \begin{tabular}{ccc}
      \includegraphics[scale=0.3]{figures/chj_r2_llf_t0p1.png} \qquad &
      %\quad \includegraphics[scale=0.5]{figures/cross.png} \quad &
      \qquad \includegraphics[scale=0.3]{figures/chj_r2_blended_t0p1.png}
    \end{tabular}
    }
  \caption{CHJ with boundary conditions given by \eqref{bcs_chj} and \eqref{bcs_chj_r2}. 
    We consider method \eqref{second-order_via_fluct} with different Riemann solvers.
    In all cases we show the Schlieren plot for the water height $h$.
    The inset figures show the momentum in the radial direction.}
\end{figure}

\begin{figure}[!h]
  \centering
    \includegraphics[scale=0.3]{figures/chj_r2_rusanov_t5p0.png} \quad
    %\includegraphics[scale=0.5]{figures/cross.png} \quad
    \quad \includegraphics[scale=0.3]{figures/chj_r2_blended_t5p0.png}
  \caption{CHJ at $t=5$ with boundary conditions given by \eqref{bcs_chj} and \eqref{bcs_chj_r2}. 
    We consider method \eqref{second-order_via_fluct} with (left) Rusanov's
    and (right) the blended solvers. 
    In all cases we superimpose the Schlieren plot for the water height $h$
    an the magnitude of the momentum with a uniform (white-to-blue) colormap. 
    The scale for colormap is (white) $0.45$ to (blue) $4.5$, 
    which corresponds to the minimum and maximum values at $t=0$.
    The inset figure shows the momentum in the radial direction.
    \label{fig:chj_r2_later}}
\end{figure}

\subsubsection{Regime II with random perturbations}\label{sec:regime_iiwp}
Given the symmetry of the grid and initial data in the section above,
the non-rotationally-symmetric perturbations seen when using the blended
solver above must arise due to the influence of numerical errors.
In order to understand whether these are evidence of a true instability,
we now introduce a non-symmetric perturbation at the inflow boundary.
We use the same mean values \eqref{bcs_chj_r2}, but now we set
\begin{subequations}
  \begin{align}
    h(x,y,t)=\tilde h_\jet, \quad
    u(x,y,t)=|\tilde \bfu_\jet| \left(\frac{x}{r_\jet}\right), \quad
    v(x,y,t)=|\tilde \bfu_\jet| \left(\frac{y}{r_\jet}\right), \qquad r=r_\jet,
  \end{align}
\end{subequations}
with $\tilde h_\jet = \frac{h_\jet}{1+\epsilon(x,y,t)}$ and $|\tilde\bfu_\jet|=|\bfu_\jet|(1+\epsilon(x,y,t))$.
Here $\epsilon(x,y,t)$ is chosen at each inflow boundary ghost cell and at each
time step as an i.i.d. random variable from the uniform distribution $[-0.01,0.01]$.
In this case, due to the large Froude number, the random perturbations in the inflow
may trigger physical instabilities, which we want to preserve. At the same time, we
do not want the solution to develop carbuncle instabilities.
In figures \ref{fig:chj_r2wp_rusanov} and \ref{fig:chj_r2wp_blended}, we show the solution
at different times considering method \eqref{second-order_via_fluct} with Rusanov's and
the blended Riemann solvers, respectively.
In the right panel of Figure \ref{fig:chj_r2wp_blended}, we show an inset figure 
with the momentum in the radial direction. Note that even though the solution is highly unstable 
(with the blended Riemann solver), no visible carbuncle instabilities are developed.

@@

  \subfloat[Blended Riemann solver\label{fig:chj_r2wp_blended}]{
    \begin{tabular}{cccc}
      \includegraphics[scale=0.24]{figures/chj_r2wp_blended_t2p0.png} &
      \includegraphics[scale=0.24]{figures/chj_r2wp_blended_t3p0.png} &
      \includegraphics[scale=0.24]{figures/chj_r2wp_blended_t4p0.png} &
      \includegraphics[scale=0.24]{figures/chj_r2wp_blended_t5p0.png}
    \end{tabular}
  }
  \caption{CHJ at different times with a random perturbation at the inflow boundary.
    We consider method \eqref{second-order_via_fluct} with different Riemann solvers. 
    In all cases we show the Schlieren plot for the water height $h$.
    The inset figure shows the momentum in the radial direction.}
\end{figure}
